Scaling through abstractions – high-performance vectorial wave simulations for seismic inversion with Devito

### 2020-06-01 paper.tex
add refs and FIXME

\textcolor{red}{It should be emphasized that unlike in,
e.g., earthquake simulation models \cite{fu201718} (more?), where
a single shot is computed, wave-equation based
geophysical inversion not only requires
fast and accurate solvers but also adavnced algorithms for
non-convex optimization of large scale constrained inverse problems. The 
high-level interface provided by \devito enables research and developement for these 
specific problem in a framework that scales to industry size problem 
automatically through its compiler. \devito has been used previously
for the solution of the wave equation within imaging frameworks including
FWI \cite{witte2018alf}, compressived least-square RTM \cite{witte2018cls} and
time-domain Wavefield Reconstruction Inversion (WRI)~\cite{rizzuti2020EAGEtwri, 
louboutin2020SEGtwri}, the later of which takles the cycle-skipping issue
present in seismic inversion.}

@@

In this section we illustrate the \devito language by demonstrating
an implementation of the acoustic wave-equation in isotropic media
%
\begin{equation}
\begin{cases}
 m \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
 u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
 d(t, .) = u(t, xr).
 \end{cases}
\label{acou}
\end{equation}
%
\textcolor{red}{FIXME: Define the terms appearing in the above equation.}
 The core of the Devito
symbolic API consists of three classes:

@@

We here provide a succinct description of distributed-memory parallelism
in Devito; the interested reader should refer to the MPI tutorial at
\href{https://github.com/devitocodes/devito/blob/v4.2/examples/mpi/overview.ipynb}{mpi-notebook}
for thorough explanations and practical examples.

@@

Earlier performance benchmarks mainly focused on roofline model analysis
\textcolor{red}{[TODO: Add reference]}.
In this study, for completeness, the runtime of \devito is therefore
compared to that of the open source hand-coded propagator
\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc}. This propagator,
described in \cite{thorbecke}, is a state of the art elastic kernel
(Equation~\ref{elas1}) and the comparisons presented here were carried out
in collaboration with its author.
\textcolor{red}{It should be emphasized that direct model comparison in
the area of exploration geophysics is challenging. The majority of production
codes are either proprioty or are simply not publicly available. Further,
\devito should not be compared against pure stencil frameworks,
such as YASK \cite{yask}, since they are unable to compute the sparse operations
required for source injection and receiver interpolation (and are therefore not
suitable for use in measurement based inversion problems). The comparison carried out
here is conducted in two dimensions. Whilst large scale 3D scaling comparisons are theoretically
possible, compute time on the Cloud is expensive. Such a comparison is not feasible at this
time and will therefore be left for future work.}

>>>>

\textcolor{red}{It should be emphasized that unlike in,
e.g., earthquake simulation models \cite{fu201718}, where
a single shot is computed, wave-equation based
geophysical inversion not only requires
fast and accurate solvers but also adavnced algorithms for
non-convex optimization of large scale constrained inverse problems. The 
high-level interface provided by \devito enables research and developement for these 
specific problem in a framework that scales to industry size problem 
automatically through its compiler. \devito has been used previously
for the solution of the wave equation within imaging frameworks including
FWI \cite{witte2018alf}, compressived least-square RTM \cite{witte2018cls} and
time-domain Wavefield Reconstruction Inversion (WRI)~\cite{rizzuti2020EAGEtwri, 
louboutin2020SEGtwri}, the later of which takles the cycle-skipping issue
present in seismic inversion.}

@@

In this section we illustrate the \devito language by demonstrating
an implementation of the acoustic wave-equation in isotropic media
%
\begin{equation}
\begin{cases}
 m(x) \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
 u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
 d(t, .) = u(t, xr).
 \end{cases}
\label{acou}
\end{equation}
%
\textcolor{red}{where $u(t, x)$ is the pressure wavefield, $xs$ is the point source position, 
$q(t)$ is thes ource time signature, $d(t, .)$ is the measured data at positions $xr$ and $m(x)$ is the 
squared slowness.}
 The core of the Devito
symbolic API consists of three classes:

@@

We here provide a succinct description of distributed-memory parallelism
in Devito; the interested reader should refer to the MPI tutorial at
\href{https://github.com/devitocodes/devito/blob/v4.2/examples/mpi/overview.ipynb}{mpi-notebook}
for thorough explanations and practical examples.
\textcolor{red}{We highlugh the main feature of the distributed framework tha enable the industry scale
applications we describe in this paper. A more thorough study of its implementation, including `mpi4numpy' and
a complete scaling analysis will be presented in a later paper. Initial results assciated with early version of \devito's
distributed computing capabilities can be found in \cite{luporini2019adp} on a conventional cluster and 
in \cite{witte2019ecl} that contains results on AWS.}

@@

Earlier performance benchmarks mainly focused on roofline model analysis
\textcolor{red}{\cite{sevendwarfs, patterson, williams2009roofline, louboutin2016ppf}}.
In this study, for completeness, the runtime of \devito is therefore
compared to that of the open source hand-coded propagator
\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc}. This propagator,
described in \cite{thorbecke}, is a state of the art elastic kernel
(Equation~\ref{elas1}) and the comparisons presented here were carried out
in collaboration with its author.
\textcolor{red}{It should be emphasized that direct model comparison in
the area of exploration geophysics is challenging. The majority of production
codes are either proprioty or are simply not publicly available. Further,
\devito should not be compared against pure stencil frameworks,
such as YASK \cite{yask}, since they are unable to compute the sparse operations
required for source injection and receiver interpolation (and are therefore not
suitable for use in measurement based inversion problems). The comparison carried out
here is conducted in two dimensions. Whilst large scale 3D scaling comparisons are theoretically
possible, compute time on the Cloud is expensive. Such a comparison is not feasible at this
time and will therefore be left for future work.}
